---
title: "Fabricated Data Preparaion"
author: "Ben Jaques-Leslie"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  project: 'MFIP workers'
  data_file: "MFIP-workers-brief-prepared-2022-12-23.RData" 
  number_ids: 5000
---

<!-- (using parameters in Rmarkdown; don't forget the r) -->
<!-- (This statement above all tabs) -->
This is my report data preparation report for `r paste(params$project)`.

<!-- (making tabs) -->
<!-- # header 1, all tabs need to be ## header 2 to nest -->

# {.tabset .tabset-pills}

## Some Info {.tabset .tabset-pills}

Click buttons below to find details about this data extraction.

### Set up

```{r setup, echo = FALSE,message = FALSE,warning = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = TRUE,
                      message = TRUE)
tictoc::tic()
library(tidyverse)
library(lubridate)
library(eaesdrrr)
library(ggplot2)
library(MNColorrrs)
library(ggthemes)
library(ggrepel)
library(ggalluvial)
library(english)
library(DBI)
library(tigris)
library(sf)
library(DBI)
library(gtsummary)
library(janitor)
library(officer)
library(officedown)
library(flextable)
library(skimr)
library(rstatix)
library(arsenal)
library(gdata)
```

```{r load mfip workers data}
load(here::here("data",params$data_file))
```

## Creating fabriated data {.tabset .tabset-pills}

### Fabricating functions

Written by Joe Ritter.

This function takes a categorical vector and generates a random distribution with the same probabilities.

```{r}
fakeCat <- function(v, n) {
    # v is integer, string, or factor.
    # n is the desired length of the fake vector.
    w <- as.character(v)
    dist <- prop.table(table(w, useNA = 'always'))
    vals <- names(dist)
    fake.v <- sample(vals, n, prob=dist, replace=TRUE)
    if (is.factor(v)) fake.v <- factor(fake.v)
    if (is.integer(v)) fake.v <- as.integer(fake.v)
    return(fake.v)
}
```

This function creates a vector of fake dates.

```{r}
fakeDate <- function(v, n)
{
  lubridate::ymd(fakeCat(v,n))
}
```


This function takes a continuous vector and generates a distribution with the same distribution.

```{r}
fakeCont <- function(v, n) {
  # Remove nulls from v
     v <- na.omit(v)
 
    # v is a variable for which a kernel density estimate makes sense.
    # n is the desired length of the fake vector.
    dist <- density(v)
    # The simulation follows one of the examples for density().
    # The default kernel is gaussian, which is why rnorm is used.  
    # Bandwidth is the SD of the # kernel.
    fake.v <- rnorm(n, mean=sample(v, n, replace=TRUE), sd=dist$bw)
    return(fake.v)                    
}
```

```{r}
fakeCont(v = cohort$Wages,n = 10)
```


```{r}
unique(cohort$MF) %>% length()

c(1:10) %>% length()

class(cohort$race)

all.equal(cohort$Wages, as.integer(cohort$Wages))

ifelse(all.equal(round(cohort$case_size_all),cohort$case_size_all) == TRUE,TRUE,FALSE)
```


```{r}
helper_unique <- function(in_original_data,in_variable, n)
{
  v <- 
    in_original_data %>% 
    select(PersonID,{{in_variable}}) %>% 
    distinct() %>% 
    pull({{in_variable}})
  
  data_type <- class(v)
  
  if(unique(v) %>% length() <= 3 & data_type != "Date")
  {
    data_type <- "character"
  }
  
  
  if(unique(v) %>% length() > 3)
  {
    data_type <- class(v)
  }
  
  if(data_type == "Date")
  {
     out <- fakeDate(
  v = v,
  n = n)
  }
  
  if(data_type == "factor")
  {
    out <-  fakeCat(
  v = v,
  n = n)
  }
    
    if(data_type == "character")
  {
    out <-  fakeCat(
  v = v,
  n = n)
    }
    
     if(data_type == "numeric")
  {
     out <- fakeCont(
  v = v,
  n = n)
     }
  
    if(data_type == "integer")
  {
     out <- fakeCat(
  v = v,
  n = n)
     }
  
  return(out)
}
```

```{r}
helper_unique(cohort,MF,n = params$number_ids) %>% table()
```

```{r}
fakeCat(cohort$race,n = 10)
```

### What needs to be unique? 

The following variables define a row in the final data set:

1.  Person identifier
2.  Baseline quarter
3   Relative quarter

Start by creating a dataframe of unique ids.

```{r}
prep_01 <- 
  tibble(PersonID = c(1:params$number_ids)) %>% 
  mutate(
    PersonID = str_pad(PersonID,width = 8, pad = "0", side = "left")
  )

prep_01 %>% skim()
```

Now we create baseline quarters off of the original data, but completely fabricated. Each PersonID has the a baseline_quarter and the distribution matches that of the larger data.

```{r}
prep_01 <- 
  prep_01 %>% 
  mutate(
   baseline_quarter = helper_unique(cohort,
                                    baseline_quarter,
                                    n = params$number_ids)
  )

prep_01 %>% skim()
```

### Adding other time-invarient factors {.tabset .tabset-pills}

Several factors in the data should not change from quarter to quarter. They include:

-   DateOfBirth
-   Gender
-   raceallcodes
-   HispanicLatinoInd
-   DeathDate
-   LanugageCode
-   WrittenLanguageCode
-   immigrant
-   age_q_start
-   age_q_end
-   race

#### Adding dates of birth and death

```{r}
prep_02 <- 
  prep_01 %>% 
  mutate(
    DateOfBirth = 
      helper_unique(cohort,
                    DateOfBirth,
                    n = params$number_ids),
    DeathDate = 
      helper_unique(cohort,
                    DeathDate,
                    n = params$number_ids)
  )

prep_02 %>% skim()
```

#### Adding other features

-   Gender
-   raceallcodes
-   HispanicLatinoInd
-   LanugageCode
-   WrittenLanguageCode
-   immigrant
-   age_q_start
-   age_q_end
-   race

```{r}
prep_02 <-
  prep_02 %>% 
  mutate(
    Gender = 
      helper_unique(cohort,
                    Gender,
                    n = params$number_ids),
    raceallcodes = 
      helper_unique(cohort,
                    raceallcodes,
                    n = params$number_ids),
    HispanicLatinoInd = 
      helper_unique(cohort,
                    HispanicLatinoInd,
                    n = params$number_ids),
    LanguageCode = 
      helper_unique(cohort,
                    LanguageCode,
                    n = params$number_ids),
    WrittenLanguageCode = 
      helper_unique(cohort,
                    WrittenLanguageCode,
                    n = params$number_ids),
    immigrant =
      helper_unique(cohort,
                    immigrant,
                    n = params$number_ids),
    age_q_start =
      helper_unique(cohort,
                    age_q_start,
                    n = params$number_ids),
    age_q_end =
      helper_unique(cohort,
                    age_q_end,
                    n = params$number_ids),
    race =
      helper_unique(cohort,
                    race,
                    n = params$number_ids)
  )
  

prep_02 %>% skim()
```

### Creating panel structure

Now we duplicated these for each of the relative quarters.

```{r}
prep_03 <- 
  cohort %>% 
  select(relative_quarter) %>%
  distinct() %>% 
  arrange(relative_quarter) %>% 
  pull() %>% 
map_dfr(~
  mutate(prep_02,
    relative_quarter = .
  ))

prep_03 %>% skim()
```

### Adding time-varying features {.tabset .tabset-pills}

Find the names of columns in `cohort` but not in `prep_03` e.g. variables of which we might need to make fake versions.

```{r}
names_cohort_not_prep_03 <- 
  names(cohort)[!(names(cohort) %in% names(prep_03))]

names_cohort_not_prep_03
```

Create a list of columns from this list of which we don't need to make fake versions.

```{r}
columns_to_remove <- 
  c(
    "SocialSecurityNbr",
    "SSN",
    "personid_ui",
    "CalendarYear",
    "Quarter",
    "EmployerAccountNumber",
    "EmployerUnitNumber",
    "OfficerFlag",
    "UsedInClaimFlag",
    "WeeksWorked",
    "Parent_Cd",
    "Description",
    "qtr_end",
    "qtr_start",
    "CaseNumber",
    "quarter_end"     
)
```

Remove the above to create a list of variables that will be fabricated.

```{r}
column_to_make_fake <- 
  names_cohort_not_prep_03[! names_cohort_not_prep_03 %in% columns_to_remove]

column_to_make_fake
```

Create dataframe of fabricated data based on the columns above and apply the names of those columns to the new dataframe. 

```{r}
prep_04 <- 
  column_to_make_fake %>% 
  map_dfc(~helper_unique(cohort,
                    all_of(.),
                    n = nrow(prep_03)
                    )
  ) %>% 
  rename_with(~column_to_make_fake)

prep_04 %>% 
  skim()
```

Bind `prep_03` and `prep_04`.

```{r}
prep_xx <- 
  prep_03 %>% 
  bind_cols(prep_04)

prep_xx %>% 
  skim()
```

## Comparing original and fabricated data {.tabset .tabset-pills}

```{r}
comparedf(cohort %>% sample_n(size = 100000),prep_xx) %>% summary()
```

## Save data {.tabset .tabset-pills}

Rename fabricated data.

```{r}
mv(from = "prep_xx", to = "cohort_fabricated")
```

File name.

```{r}
file <- glue::glue("mfip-worker-fabricated-{Sys.Date()}.rdata")

file
```

Save data.

```{r}
save(cohort_fabricated,
     file = here::here(
       "output",
       "rdata",
       file
     )
     )
```

Remove all data elements.

```{r remove data, results='hide'}
rm(list=ls())
gc()

```

## Extraction time {.tabset .tabset-pills}

```{r}
tictoc::toc()
```

